Exploratory analysis of year 2 surveillance data

### 2023-02-27 code/analysis_code/exploratory_analysis_LML.qmd
exploratory analysis

```{r}
library(tidyverse)
library(here)
library(ggpmisc)

@@

```{r}
covid_combine %>% ggplot(aes(pcr_pos_7dma)) +
  geom_histogram()
summary(covid_combine$pcr_pos_7dma)
```

@@

```{r}
covid_combine %>% ggplot(aes(pcr_pos)) +
  geom_histogram()
summary(covid_combine$pcr_pos)
```

@@

```{r}
sum_daily_copies_county <- all_data_combine %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies))

  sum_daily_copies_county %>% ggplot(aes(log10(sum_total_copies))) +
  geom_histogram()

  log10_sum_daily_copies <- log10(sum_daily_copies_county$sum_total_copies)
  summary(log10_sum_daily_copies)
```

### Look at case 7 day moving average over time

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_7dma = mean(pcr_pos_7dma)) %>% 
  ggplot(aes(x=date, y=mean_7dma)) +
  geom_point() +
 stat_smooth(method = "gam")

@@

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_7dma = mean(pcr_pos_7dma),
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_7dma)) +
  geom_bar(stat = "identity", color = '#634490') +
  stat_smooth(method = "gam", color = 'black')

@@

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_pcr_pos)) +
  geom_bar(stat = "identity", color = '#634490') +
  stat_smooth(method = "gam", color = 'black') 
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  #drop_na() %>% #remove NAs from analysis
  group_by(date, facility) %>% #this will allow me to average the two targets
  summarize(mean_total_copies = mean(total_copies)) %>% #average across N1 and N2
  ungroup(facility) %>% #need to do this so I can add the avg total copies of each WRF
  summarize(sum_total_copies = sum(mean_total_copies)) %>% #total daily county copies
  ggplot(aes(date, log10(sum_total_copies))) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

  ggplot(aes(date, log10(sum_total_copies), color = target)) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
 stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

@@

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

@@

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(copy_num_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(x=date)) +
  #geom_point(aes(y=log10(sum_mean_copies_L)/mean_7dma)) + 
  geom_line(aes(y=(log10(sum_mean_copies_L)/mean_7dma), color = "red")) +
  geom_line(aes(y=log10(mean_7dma)), color = "blue")

@@

>>>>

```{r}
library(tidyverse)
library(here)
library(ggpmisc)
library(ggthemes)

@@

```{r}
covid_combine %>% ggplot(aes(pcr_pos_7dma)) +
  geom_histogram(fill = "#C06C84") +
xlab(label = "Clinical Case Seven Day Moving Average") +
ylab(label = "Count") +
theme_light()

ggsave(here("figures/exploratory/variate/7DMA_histo.png"))
summary(covid_combine$pcr_pos_7dma)
```

@@

```{r}
covid_combine %>% ggplot(aes(pcr_pos)) +
  geom_histogram(fill = '#F8B195') +
  xlab(label = "Daily Positive Clinical Cases") +
  ylab(label = "Count") +
  theme_light()

ggsave(here("figures/exploratory/variate/pos_cases_histo.png"))
summary(covid_combine$pcr_pos)
```

@@

```{r}
sum_daily_copies_county <- all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_total_copies = mean(total_copies)) %>% 
  ungroup(facility) %>% 
  summarize(sum_total_copies = sum(mean_total_copies))

  sum_daily_copies_county %>% ggplot(aes(log10(sum_total_copies), fill = target)) +
  geom_histogram() +
      scale_fill_manual(name = "Target",values = c("#F67280", "#355C7D")) +
    xlab(label = "Log10 Total Daily County-Level Copies") +
    ylab(label = "Count") +
    theme_light()

ggsave(here("figures/exploratory/variate/total_copies_histo.png"))
  log10_sum_daily_copies <- log10(sum_daily_copies_county$sum_total_copies)
  summary(log10_sum_daily_copies)
```

@@

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_7dma = mean(pcr_pos_7dma),
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_7dma)) +
  geom_bar(stat = "identity", color = '#C06C84') +
  stat_smooth(method = "gam", color = 'black') +
  xlab(label = "Date") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/7DMA_time_plot.png"))

@@

```{r}
all_data_combine %>% group_by(date) %>% 
  summarize(mean_pcr_pos = mean(pcr_pos)) %>% 
  ggplot(aes(x = date, y = mean_pcr_pos)) +
  geom_bar(stat = "identity", color = '#F8B195') +
  stat_smooth(method = "gam", color = 'black') +
  ylab(label = "Positive Cases Per Day") +
  xlab(label = "Date") +
  theme_light()

ggsave(here("figures/exploratory/covariate/pos_cases_time.png"))
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na(target) %>% #remove NAs from analysis
  group_by(date, facility) %>% #this will allow me to average the two targets
  summarize(mean_total_copies = mean(total_copies)) %>% #average across N1 and N2
  ungroup(facility) %>% #need to do this so I can add the avg total copies of each WRF
  summarize(sum_total_copies = sum(mean_total_copies)) %>% #total daily county copies
  ggplot(aes(date, log10(sum_total_copies))) +
  geom_point(color = "#80256D") +
  stat_smooth(method = "gam", color = "#80256D") +
   xlab(label = "Date") +
  ylab(label = "Log10 Total Daily County-Level Copies") +
  theme_light()

ggsave(here("figures/exploratory/covariate/n1_n2_avg_time.png"))
```

@@

  ggplot(aes(date, log10(sum_total_copies), color = target)) +
  geom_point() +
  stat_smooth(method = "gam") +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Date") +
  ylab(label = "Log10 Total Daily County-Level Copies") +
  theme_light()

 ggsave(here("figures/exploratory/covariate/n1_n2_time.png"))
```

@@

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L))) +
  geom_point() +
  stat_smooth(method = "gam")
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L)) %>% 
  ggplot(aes(date, log10(sum_mean_copies_L),color = target)) +
 stat_smooth(method = "gam") +
  geom_point()

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

@@

```{r}
all_data_combine %>% filter(target == "N1") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

@@

```{r}
all_data_combine %>% filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies_L), mean_7dma)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq()
```

### Look at log10 total daily copies per 7DMA

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies), mean_7dma)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq() +
  xlab(label = "Log10 Total Daily County-Level Copies") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_avg_vs_7DMA.png"))
```

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(log10(sum_mean_copies), mean_7dma, color = target)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Log10 Total Daily County-Level Copies") +
  ylab(label = "Clinical Case Seven Day Moving Average") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_vs_7DMA.png"))
```

@@

```{r}
all_data_combine %>% #filter(target == "N2") %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_copies_L = mean(avg_copy_L), 
            mean_7dma = mean(pcr_pos_7dma)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies_L = sum(mean_copies_L),
            mean_7dma = mean(mean_7dma)) %>% 
  ggplot(aes(x=date)) +
  #geom_point(aes(y=log10(sum_mean_copies_L)/mean_7dma)) + 
  geom_line(aes(y=(log10(sum_mean_copies_L)/mean_7dma), color = "red")) +
  geom_line(aes(y=log10(mean_7dma)), color = "blue")

@@

### Log10 Daily copies N1 & N2 vs daily positive cases

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, target, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_pcr_pos = mean_pcr_pos) %>% 
  ggplot(aes(log10(sum_mean_copies), mean_pcr_pos, color = target)) +
  geom_point() +
  stat_poly_line() +
  stat_poly_eq() +
  scale_color_manual(name = "Target",values = c("#F67280", "#355C7D")) +
  xlab(label = "Log10 Total Daily County-Level Copies") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_vs_pcr_pos.png"))
```

### Log10 Daily copies average vs daily positive cases

```{r}
all_data_combine %>% 
  drop_na() %>% 
  group_by(date, facility) %>% 
  summarize(mean_daily_copies = mean(total_copies), 
            mean_pcr_pos = mean(pcr_pos)) %>% 
  ungroup(facility) %>% 
  summarize(sum_mean_copies = sum(mean_daily_copies),
            mean_pcr_pos = mean_pcr_pos) %>% 
  ggplot(aes(log10(sum_mean_copies), mean_pcr_pos)) +
  geom_point(color = "#80256D") +
  stat_poly_line(color = "#80256D") +
  stat_poly_eq() +
  xlab(label = "Log10 Total Daily County-Level Copies") +
  ylab(label = "Daily Positive Clinical Cases") +
  theme_light()

ggsave(here("figures/exploratory/covariate/total_copies_avg_vs_pcr_pos.png"))
```
